---
title: "Descriptive statistics of STEPWAT2 biomass data"
author: "Martin Holdrege"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

The purpose of this file is to create tables of descriptive statistics
that provide numbers that are useful when writing the results of the manuscript,
so that the underlying code doesn't need to be run everytime you want to look
up a number

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dependencies

```{r packages, message = FALSE, warning=FALSE,}
library(tidyverse)
library(knitr)
```

The 02_summarize_bio.R script creates the dataframes used below.

```{r source, message = FALSE, warning=FALSE, results='hide', cache=TRUE}
# note set cache=TRUE to cache=FALSE once done developing this code,
# because it takes disk space
source("scripts/02_summarize_bio.R")
```

# Grazing effects


```{r}
pft5_d_wgcm_heavy %>% 
  filter(c4 == 'c4on') %>% 
  group_by(graze, PFT) %>% 
  summarize(mean = round(mean(bio_diff), 1),
            median = round(median(bio_diff), 1)) %>% 
  kable(caption = 'Mean and median scaled percent changes in biomass 
        when transitioning from heavy grazing to another grazing level.
        Values are averages/medians across all climate scenarios.')
  
```

Percent of sites that had more/less biomass when changing from heavy grazing
to another intensity. 

```{r}
pft5_d_wgcm_heavy %>% 
  filter(c4 == 'c4on') %>% 
  group_by(graze, PFT) %>% 
  summarize(percent_increase = mean(bio_diff > 0)*100,
            percent_decrease = 100 - percent_increase) %>% 
  kable(caption = "Percent of sites that had biomass increasses/decreases when
        changing from heavy grazing, average across all climate scenarios")
```

# Climate effects


