---
title: "Biomass models"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Dependencies

Sourcing the script that creates the needed dataframes.

```{r message = FALSE, warning = FALSE}
# note--knit directory set to project directory
source("scripts/02_summarize_bio.R")
library(knitr)
library(lme4)
```

# Data prep


```{r}

bio_a <- pft5_bio2 %>% 
  filter(c4 == "c4on") %>%  # for now just fitting models with c4on simulation data
  mutate(RCP_years = paste(RCP, years, sep = "_")) %>% 
  split(.$PFT) # separate df for each PFT


```

# Models

## LM

Fitting OLS linear models (ANOVAs), not accounting for repeated measures
of the same plot.


```{r}
form_lm <- as.formula("biomass ~ RCP_years*graze")

# fit lm for each PFT
lm_l <- map(bio_a, function(df) {
  lm(formula = form_lm, data = df)
})

```

Putting together the model table

```{r}

# combining anova tables (made tidy)
df_lm1 <- map_dfr(lm_l, 
                 function(x) broom::tidy(anova(x)),
                 .id = "PFT")

df_lm2 <- df_lm1 %>% 
  group_by(PFT) %>% 
  mutate(SST = sum(sumsq), # total sum of squares
         # proportion of total sum of squares for that term
         `SS/SST` = sumsq/SST,
         SSM = SST - sumsq[term == "Residuals"], # SS model
         `SS/SSM` = sumsq/SSM, # proportion of SSM for that variable
         `SS/SSM` = ifelse(term == "Residuals", NA, `SS/SSM`)) %>% 
  select(-SST, - SSM)

round_vec <- c(0, 0, 0, 0, 0, 0, 2, 3, 3) # how much to round each col

kable(df_lm2,
     digits = round_vec,
     caption = "Results from ANOVAs for each PFT, showing proportion of
     SST contributed by each term, and proportion of SSM contributed by 
     each term.")

df_lm2 %>% 
  filter(term == "RCP_years:graze") %>% 
  arrange(desc(`SS/SSM`)) %>% 
  kable(digits = round_vec,
        caption = "Models ordered by relative importance of the interaction term")
  
```


## LMM

Linear mixed models where site is a random effect (just a random intercept for
now)

Fit the model 

```{r}
form_lmm <- as.formula("biomass ~ RCP_years*graze + (1 | site)")

# fit lm for each PFT
lmm_l <- map(bio_a, function(df) {
  lmer(formula = form_lmm, data = df)
})


```

Create model tables. Note SS/SSM is the same as when just fitting the 
simple linear model. 

```{r warning = FALSE}

# combining anova tables (made tidy)
df_lmm1 <- map_dfr(lmm_l, 
                 function(x) broom::tidy(anova(x)),
                 .id = "PFT")

df_lmm2 <- df_lmm1 %>% 
  group_by(PFT) %>% 
  mutate(SSM = sum(sumsq), # the lmm doesn't give residual ss, so summing
         # give SSM
         # proportion of total sum of squares for that term
         `SS/SSM` = sumsq/SSM # proportion of SSM for that variable
         ) %>% 
  select(-SSM)

round_vec2 <- c(0, 0, 0, 0, 0, 0, 3)

kable(df_lmm2,
     digits = round_vec2,
     caption = "Results from mixed models for each PFT, showing 
     proportion of SSM contributed by 
     each term.")

df_lmm2 %>% 
  filter(term == "RCP_years:graze") %>% 
  arrange(desc(`SS/SSM`)) %>% 
  kable(digits = round_vec2,
        caption = "Models ordered by relative importance of the interaction term")
  

```




