---
title: "Comparing AIM C4 grass data to upscaled STEPWAT2 ouput"
author: "Martin Holdrege"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

The purpose of this analysis is to plot AIM data showing observed 
presence/absence of C4 perennial
grasses, and compare that to where stepwat2 is simulating c4pgrass present when that
data is upscaled. 
Additional analyses are also be done to look at AIM sites that show C4grass
but for which the upscaled data doesn't simulate c4 grass, and then look
at what c4 grass species occur there. Also just tally for how many aim sites
c4 grass presence/absence is misspecified

# Dependencies

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(terra)
library(knitr)
source("src/mapping_functions.R") # knit directory is project directory
source("src/fig_params.R")
```

# Read in data 

### Rasters

Upscaled stepwat2 output (created in the 03_interpolate.R script),
for c4 grasses under current conditions, light grazing.

```{r }
C4_r <- rast(file.path("data_processed/interpolated_rasters/biomass",
               "c4on_C4Pgrass_biomass_Current_Current_Light_Current.tif"))
```

### AIM

2011-2015 AIM vegetation data that Rachel Renne has compiled

```{r }
aim1 <- read_csv("data_raw/AIM/LMF_ALL.data.2011-2015.FINAL_v2.csv")
```

### Site locations

Location of the 200 sites for which simulations were run,
and whether C4pgrass is simulated to be present there under current 
conditions. 

File created in "03_interpolate.R" script

```{r }
sites_c4 <- read_csv(
  "data_processed/site-num_C4Pgrass-presence_c4off.csv")
```

### Ecoregion

Sshapefile of level 3 ecoregions , downloaded from here:
https://gaftp.epa.gov/EPADataCommons/ORD/Ecoregions/us/us_eco_l3.zip

Useful for grouping summaries of the AIM data by region

```{r }
eco1 <- vect("data_raw/ecoregion_level_3/us_eco_l3.shp")
```

# Data processing 

## AIM

the WBG and WRG columns are c4 perennial bunch grasses and c4 perennial 
rhizomatous grasses, respectively. 

```{r }
aim2 <- aim1 %>% 
  select(PLOTKEY, latitude, longitude, WBG, WRG) %>% 
  #' are C4 P grasses present?
  mutate(C4Pgrass = ifelse(WBG > 0 | WRG > 0, "present", "absent")) 

mean(aim2$C4Pgrass == "present") * 100 #' % of sites w/ c4pgrass present
```

Creating a SpatVector (terra package)
for some reason logical data column gets converted to character (can't try
to force it to be logical)

```{r }
aim_sv1 <- vect(aim2, geom = c("longitude", "latitude"),
                crs = "EPSG:4326") #' crs is wgs 1984, lat/lon
```

## sites

creating SpatVector

```{r }
sites_c4_sv1 <- vect(sites_c4, geom = c("x", "y"),
                     crs = "EPSG:4326")
```

## Extract data for each AIM site

Extracting up-scaled stepwat2 biomass values for each AIM site

```{r warning = FALSE}
aim_sv2 <- aim_sv1

# The 2nd column contains values extracted from the raster for each AIM site

aim_sv2$C4Pgrass_biomass <- extract(C4_r, aim_sv1)[, 2]
aim_sv2$C4Pgrass_aim <- aim_sv2$C4Pgrass #' presence/absence for aim data
aim_sv2$C4Pgrass <- NULL

# creating column of presence/absence of up-scaled data
aim_sv2$C4Pgrass_mod <- ifelse(aim_sv2$C4Pgrass_biomass > 0, "present", "absent")
```

Adding in the eco-region

```{r}


 # reproject to same projection as eco1 (albers equal area)
aim_sv3 <- project(aim_sv2, eco1)

# plot(eco1)
# points(aim_sv_albers1)

# eco regions of each of the AIM points
eco_aim <- extract(eco1, aim_sv3)

# adding eco region columns
aim_sv3 <- cbind(aim_sv3, eco_aim[, c("US_L3CODE", "US_L3NAME")])


```


CONTINUE HERE:
TODO--create map that shows the eco regions
within each eco region create a contingency table
and extract which species have been present in sites where no c4 species
were simulated, and how many times they've been absent. 

# Summarise AIM data

Contingency table showing C4pgrass presence absence in observed (aim data),
and modeled (up-scaled STEPWAT2)

```{r warning = FALSE}
table(as.data.frame(aim_sv2[, c("C4Pgrass_aim", "C4Pgrass_mod")])) 
```




# MAPs

map(s) showing AIM C4Pgrass occurrence on top of interpolated STEPWAT2 
C4Pgrass biomass data, as well as the 200 simulation sites
base of the map for several panels

```{r, error=TRUE, results='hide', fig.show='hide'}
base_map <- function(title) {
  image_bio(C4_r, subset = 1, title = title)
}
# due to some unknown issue this throws an error
# the first time it is run (when image resolution is high), so calling
# function here where it can't cause harm
# (see https://github.com/rspatial/terra/issues/30)
safely(base_map(title = "Current C4Pgrass, light grazing"))

```


```{r warning = FALSE}

# for some reason the code is breaking if I don't recreate aim_sv1
# again here
aim_sv1 <- vect(aim2, geom = c("longitude", "latitude"),
                crs = "4326") #' crs is wgs 1984, lat/lon

pdf("figures/biomass_maps/C4Pgrass-bio_AIMC4-Pgrass-presence.pdf",
    width = wfig6, height = hfig6)

pch1 <- 16
cex1 <- 0.4

par(mar = mar, mgp = mgp)
layout(layout.matrix6, widths = widths6, heights = heights6)

base_map(title = "Current C4Pgrass, light grazing")


# AIM plots on top


base_map(title = "AIM plots where C4Pgrass absent (black points)")
plot(subset(aim_sv1, subset = aim_sv1$C4Pgrass == "absent"), 
     col = cols_c4present["absent"], 
     add = TRUE, pch = pch1, cex = cex1)

base_map(title = "AIM plots where C4Pgrass present (blue points)")
plot(subset(aim_sv1, subset = aim_sv1$C4Pgrass == "present"), 
     col = cols_c4present["present"], 
     add = TRUE, pch = pch1, cex = cex1)

base_map(title = "All AIM plots")
plot(aim_sv1, 
     col = cols_c4present[aim_sv1$C4Pgrass], 
     add = TRUE, pch = pch1, cex = cex1/2)


# showing 200 sites

base_map(title = "200 simulation sites (blue = C4Pgrass presence")
plot(sites_c4_sv1, 
     col = cols_c4present[sites_c4_sv1$C4Pgrass], 
     add = TRUE, pch = pch1, cex = 0.8)

dev.off()
```

