---
title: "Comparing AIM C4 grass data to upscaled STEPWAT2 ouput"
author: "Martin Holdrege"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

The purpose of this analysis is to plot AIM data showing observed 
presence/absence of C4 perennial
grasses, and compare that to where stepwat2 is simulating c4pgrass present when that
data is upscaled. 
Also just tally for how many aim sites
c4 grass presence/absence is mis-specified, by eco-region.

**TO DO**: Additional analyses are need to be done to look at AIM sites that show C4grass
but for which the upscaled data doesn't simulate c4 grass, and then look
at what C4 grass species occur there.

# Dependencies

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(terra)
library(knitr)
# library(sf)
source("src/mapping_functions.R") # knit directory is project directory
source("src/fig_params.R")
source("src/general_functions.R")
```

# Read in data 

### Rasters

Upscaled stepwat2 output (created in the 03_interpolate.R script),
for c4 grasses under current conditions, light grazing.

```{r }
C4_r <- rast(file.path("data_processed/interpolated_rasters/biomass",
               "c4on_C4Pgrass_biomass_Current_Current_Light_Current.tif"))
```

### AIM

2011-2015 AIM vegetation data that Rachel Renne has compiled

```{r }
aim1 <- read_csv("data_raw/AIM/LMF_ALL.data.2011-2015.FINAL_v2.csv",
                 show_col_types = FALSE)
```

cover of individual C4 grasses (Also compiled by RR).

```{r}
aimc4_1 <- read_csv("data_raw/AIM/LMF_c4grasses.csv",
                 show_col_types = FALSE)
```

### Site locations

Location of the 200 sites for which simulations were run,
and whether C4pgrass is simulated to be present there under current 
conditions. 

File created in "03_interpolate.R" script

```{r }
sites_c4 <- read_csv(
  "data_processed/site-num_C4Pgrass-presence_c4off.csv", show_col_types = FALSE)
```

### Ecoregion

Shapefile of level 3 ecoregions , downloaded from here:
https://gaftp.epa.gov/EPADataCommons/ORD/Ecoregions/us/us_eco_l3.zip

Useful for grouping summaries of the AIM data by region

```{r }
eco1 <- vect("data_raw/ecoregion_level_3/us_eco_l3.shp")
```

# Data processing 

## AIM

the WBG and WRG columns are c4 perennial bunch grasses and c4 perennial 
rhizomatous grasses, respectively. 

```{r }
aim2 <- aim1 %>% 
  select(PLOTKEY, latitude, longitude, WBG, WRG) %>% 
  #' are C4 P grasses present?
  mutate(C4Pgrass = ifelse(WBG > 0 | WRG > 0, "present", "absent")) 

mean(aim2$C4Pgrass == "present") * 100 #' % of sites w/ c4pgrass present
```
Keeping only perennial c4 species

```{r}
aimc4_2 <- aimc4_1 %>% 
  filter(Duration != "Annual")
```

running a check between the to AIM files

```{r}

if(!all(aimc4_2$PLOTKEY %in% aim2$PLOTKEY[aim2$C4Pgrass == "present"])) {
  stop('some plots that are present int c4 grass data set are not 
  present in the overall dataset')
}

if (any(aimc4_2$PLOTKEY %in% aim2$PLOTKEY[aim2$C4Pgrass == "absent"])) {
  stop('some sites are marked as c4 grass being present, but c4 species recorded')
}

```


Creating a SpatVector (terra package)
for some reason logical data column gets converted to character (can't 
force it to be logical)

```{r }
aim_sv1 <- vect(aim2, geom = c("longitude", "latitude"),
                crs = "EPSG:4326") #' crs is wgs 1984, lat/lon
```

## sites

creating SpatVector

```{r }
sites_c4_sv1 <- vect(sites_c4, geom = c("x", "y"),
                     crs = "EPSG:4326")
```

## Extract data for each AIM site

Extracting up-scaled stepwat2 biomass values for each AIM site

```{r warning = FALSE}
aim_sv2 <- aim_sv1

# The 2nd column contains values extracted from the raster for each AIM site

aim_sv2$C4Pgrass_biomass <- extract(C4_r, aim_sv1)[, 2]
aim_sv2$C4Pgrass_aim <- aim_sv2$C4Pgrass #' presence/absence for aim data
aim_sv2$C4Pgrass <- NULL

# creating column of presence/absence of up-scaled data
aim_sv2$C4Pgrass_mod <- ifelse(aim_sv2$C4Pgrass_biomass > 0, "present", "absent")
```

Adding in the eco-region. Map of the ecoregions can be seen 
[here](https://gaftp.epa.gov/EPADataCommons/ORD/Ecoregions/us/Eco_Level_III_US.pdf)

```{r}

 # reproject to same projection as eco1 (albers equal area)
aim_sv3 <- project(aim_sv2, eco1)

# plot(eco1)
# points(aim_sv_albers1)

# eco regions of each of the AIM points
eco_aim <- extract(eco1, aim_sv3)

# adding eco region columns
aim_sv3 <- cbind(aim_sv3, eco_aim[, c("US_L3CODE", "US_L3NAME")])

# if want to crop ecoregion polygons
# eco_crop <- crop(eco1, y = ext(aim_sv3))
# plot(eco_crop)

```


# Summarise AIM data

Contingency table showing C4pgrass presence absence in observed (aim data),
and modeled (up-scaled STEPWAT2) for all AIM plots

```{r warning = FALSE}
con1 <- table(as.data.frame(aim_sv3[, c("C4Pgrass_aim", "C4Pgrass_mod")]))

# percent correctly classified
round(pcc(con1))

addmargins(con1) 

proportions(con1, 2)# column percents

```

Contingency tables for each ecoregion, with column and row sums
as well as tables showing cell percents as percent of column sum and row sum,
respectively

```{r}
# separate dataframe for each eco-region
aim_l1 <- as.data.frame(aim_sv3) %>% 
  # convert to factor so contingency tables keep all levels even if 0
  mutate(C4Pgrass_aim = factor(C4Pgrass_aim), 
         C4Pgrass_mod = factor(C4Pgrass_mod)) %>% 
  split(., .$US_L3CODE, drop = TRUE)

n_aim <- map_dbl(aim_l1, nrow) %>% 
  sort(decreasing = TRUE)

# sorting list, based on how many aim sites are in each ecoregioin
aim_l1 <- aim_l1[names(n_aim)]

# list of contingency tables
con_l <- map(aim_l1, function(x) {
  table(x[, c("C4Pgrass_aim", "C4Pgrass_mod")])
})
```

Percent correctly classified by ecoregion (top number is the ecoregion,
bottom is the pcc)

```{r}
map_dbl(con_l, pcc) %>% 
  round()
```

Contingency tables

```{r}
# print contingency tables

regions <- names(con_l)
for (i in seq_along(con_l)){
  con <- con_l[[i]]
  US_L3NAME <- unique(aim_l1[[i]]$US_L3NAME) # region name
  cat(paste0("\n\nEco region:", regions[i], ", ", US_L3NAME, "\n\n"))
  
  print(addmargins(con))
  
  cat("\ncolumn proportion\n\n")
  print(round(proportions(con, 2), 2))# column proportion
  cat("\nrow proportion\n\n")
  print(round(proportions(con, 1), 2)) # row proportion
  cat("\n----------------------------")

}

con <- con_l[[1]]
addmargins(con) %>% str()
x <- addmargins(con)
pcc <- (x[1, 1] + x[2,2])/x[3,3]*100
```

## C4 grass species at sites where not modelled with STEPWAT2

```{r}

# plotkeys of aim sites where c4 grasses incorrectly
# modelled as absent
plots_c4_missed <- aim_sv3 %>% 
  as_tibble() %>% 
  filter(C4Pgrass_mod == "absent", # modeled as absent
         C4Pgrass_aim == "present") %>%  # but actually present
  pull(PLOTKEY)

aimc4_missed <- aimc4_2 %>% 
  filter(PLOTKEY %in% plots_c4_missed) %>% 
  left_join(as_tibble(aim_sv3[, c("PLOTKEY", "US_L3CODE", "US_L3NAME")]),
            by = "PLOTKEY")

c4_spp_pcent_missed <- aimc4_missed %>% 
  group_by(US_L3CODE, US_L3NAME) %>% 
   # number of plots in the region
  mutate(n_plots = length(unique(PLOTKEY))) %>%
  group_by(US_L3CODE, US_L3NAME, CPLANT) %>% 
  summarise(pcent_plots = n()/unique(n_plots)*100,
            pcent_plots = round(pcent_plots, 1),
            .groups = 'drop') %>% 
  arrange(US_L3CODE, desc(pcent_plots))
  
kable(c4_spp_pcent_missed,
      caption = paste("The percent of AIM sites where a given C4 grass species",
                       "is present, out of AIM sites where upscaled STEPWAT2 data",
                        "incorrectly classified C4Pgrass as absent"
                       ))
```


# MAPs

map(s) showing AIM C4Pgrass occurrence on top of interpolated STEPWAT2 
C4Pgrass biomass data, as well as the 200 simulation sites
base of the map for several panels

```{r, error=TRUE, results='hide', fig.show='hide'}
base_map <- function(title) {
  image_bio(C4_r, subset = 1, title = title)
}
# due to some unknown issue this throws an error
# the first time it is run (when image resolution is high), so calling
# function here where it can't cause harm
# (see https://github.com/rspatial/terra/issues/30)
safely(base_map(title = "Current C4Pgrass, light grazing"))

```


```{r warning = FALSE}

# for some reason the code is breaking if I don't recreate aim_sv1
# again here
aim_sv1 <- vect(aim2, geom = c("longitude", "latitude"),
                crs = "4326") #' crs is wgs 1984, lat/lon

pdf("figures/biomass_maps/C4Pgrass-bio_AIMC4-Pgrass-presence.pdf",
    width = wfig6, height = hfig6)

pch1 <- 16
cex1 <- 0.4

par(mar = mar, mgp = mgp)
layout(layout.matrix6, widths = widths6, heights = heights6)

base_map(title = "Current C4Pgrass, light grazing")


# AIM plots on top


base_map(title = "AIM plots where C4Pgrass absent (black points)")
plot(subset(aim_sv1, subset = aim_sv1$C4Pgrass == "absent"), 
     col = cols_c4present["absent"], 
     add = TRUE, pch = pch1, cex = cex1)

base_map(title = "AIM plots where C4Pgrass present (blue points)")
plot(subset(aim_sv1, subset = aim_sv1$C4Pgrass == "present"), 
     col = cols_c4present["present"], 
     add = TRUE, pch = pch1, cex = cex1)

base_map(title = "All AIM plots")
plot(aim_sv1, 
     col = cols_c4present[aim_sv1$C4Pgrass], 
     add = TRUE, pch = pch1, cex = cex1/2)


# showing 200 sites

base_map(title = "200 simulation sites (blue = C4Pgrass presence")
plot(sites_c4_sv1, 
     col = cols_c4present[sites_c4_sv1$C4Pgrass], 
     add = TRUE, pch = pch1, cex = 0.8)

dev.off()
```

