---
title: "Comparing AIM C4 grass data to SOILWAT2 output"
subtitle: "Comparing PL96 vs PL96&TS76 SOILWAT2 implementation"
author: "Martin Holdrege"
date: "`r lubridate::today()`"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
---

Purpose: compare AIM perennial C4 grass occurrence data to
where SOILWAT2 is simulating c4pgrass present, and compare two implementations
of the equation that dictates where C4 grasses are simulated as present
(i.e. PL96 (original version) and PL96&TS76 (potential new version))


# Dependencies

```{r warning = FALSE, message = FALSE}
source("src/general_functions.R")
library(tidyverse)
library(terra)
library(sf)
library(knitr)
```

# Read in data 

### SOILWAT2 gridded data

SOILWAT2 data (c4 grass cover) from Daniel

```{r }
pl96a <- rast("data_raw/SOILWAT2/grassC4frac-PL96_yr_SOILWAT2_wUS-gm_historical_gn_19790101-20201231-clim.nc")
pl96_ts76a <- rast("data_raw/SOILWAT2/grassC4frac-PL96&TS76_yr_SOILWAT2_wUS-gm_historical_gn_19790101-20201231-clim.nc")
```

### AIM

2011-2015 AIM vegetation data that Rachel Renne has compiled

```{r }
aim1 <- read_csv("data_raw/AIM/LMF_ALL.data.2011-2015.FINAL_v2.csv",
                 show_col_types = FALSE)
```

# Data processing 

## AIM

the WBG and WRG columns are c4 perennial bunch grasses and c4 perennial 
rhizomatous grasses, respectively. 

```{r }
aim2 <- aim1 %>% 
  select(PLOTKEY, latitude, longitude, WBG, WRG) %>% 
  #' are C4 P grasses present?
  mutate(C4Pgrass = ifelse(WBG > 0 | WRG > 0, "present", "absent")) 

```

Percent of AIM sites where C4 grasses are present:

```{r}
mean(aim2$C4Pgrass == "present") * 100 #' % of sites w/ c4pgrass present
```

Creating a SpatVector (terra package)
for some reason logical data column gets converted to character (can't 
force it to be logical)

```{r }
aim_sv1 <- vect(aim2, geom = c("longitude", "latitude"),
                crs = "EPSG:4326") #' crs is wgs 1984, lat/lon
```


## Extract data for each AIM site

Extracting soilwat2 cover values for each AIM site

```{r warning = FALSE}
aim_sv2 <- aim_sv1

# The 2nd column contains values extracted from the raster for each AIM site

aim_sv2$pl96_cover <- extract(pl96a, aim_sv1)[, 2] # c4pgrass cover
aim_sv2$pl96ts76_cover <- extract(pl96_ts76a, aim_sv1)[, 2] # c4pgrass cover


aim_sf3 <- aim_sv2 %>% 
  sf::st_as_sf() %>% 
  # presence/absence for aim data
  rename(aim = C4Pgrass) %>% 
  mutate(
    # presence/absence for soilwat2 data at the aim points
    pl96 = ifelse(pl96_cover > 0, "present", "absent"),
    pl96ts76 = ifelse(pl96ts76_cover > 0, "present", "absent")
  ) %>% 
  select(-matches('cover'))

```



# Summarise AIM data

Contingency table showing C4pgrass presence absence in observed (aim data),
and modeled (SOILWAT2) for all AIM plots

```{r warning = FALSE}
aim_4 <- st_drop_geometry(aim_sf3)
con1_pl96 <- table(aim_4[, c("aim", "pl96")])
con1_pl96ts76 <- table(aim_4[, c("aim", "pl96ts76")])



```

Percent correctly classified

```{r}
round(pcc(con1_pl96))
round(pcc(con1_pl96ts76))
```

Contingency tables

```{r comment = ""}

addmargins(con1_pl96) # counts
addmargins(con1_pl96ts76) 
proportions(con1_pl96, 2) %>% 
  round(3)# column percents
proportions(con1_pl96ts76, 2) %>% 
  round(3) # column percents
```


